### 2023-11-16 ms.qmd
Updates

# Load data
	data  <-  read.csv("./data/.csv")
```

```{r, dataclean}
#| label: dataclean
#| echo: false
#| warning: false

# Remove individuals who did not participate (more than 15 NAs), remove trials [36-40] (Only in associative) and split treatment into Temp and Cort
data_associative <- data %>%
  group_by(lizard_id) %>%
    filter(sum(is.na(FC_associative)) <= 15) %>%
    filter(trial_associative <= 35) %>%
  ungroup()  %>% 
    mutate(temp = gsub("[AB]_", "", trt),
          cort = gsub("_[2][38]", "", trt))  %>% 
    mutate(group = factor(group, levels = c("R_B", "B_R"))) %>%
  data.frame()

# Standarize data by trial (i.e. make the first trial where each individual participated their trial 1)

data_associative <- data_associative %>%
  group_by(lizard_id) %>%
  mutate(
    first_non_na = min(which(!is.na(FC_associative))),
    Associative_Trial = ifelse(!is.na(first_non_na),trial_associative - first_non_na + 1, trial_associative))%>%
    filter(Associative_Trial >= 1) %>%
   ungroup() %>% data.frame()

write.csv(data_associative, file= "./output/Checking/data_associative.csv")

# Split data by species
deli_associative <- data_associative %>% filter (species == "delicata")
guich_associative <- data_associative %>% filter (species == "guichenoti")

>>>>

# Load data
	data  <-  read.csv("./data/Stressor.csv")
```

```{r, dataexplore}
#| label: dataexplore
#| echo: false
#| warning: false

# Divide Treatment in CORT and Temperature
data <- data %>%
      mutate(temp = gsub("[AB]_", "", trt),
          cort = gsub("_[2][38]", "", trt))  %>% 
  data.frame()

# MASS
# Create new database to analyse increases in mass (i.e. having mass after and before in different columns)
data_2 <- pivot_wider(data, id_cols=c(lizard,temp,cort,species,food_ingested), names_from = Day, values_from = mass)
data_2 <- data_2 %>% mutate(delta_mass=After-Before) 
data_2 <- data_2 %>% mutate(rescaled_delta_mass = delta_mass - min(delta_mass) + 1)
write.csv(data_2, "./data/Stressor_2.csv")
# Transform mass into mg in new database
data_2$mass <- data$mass*1000

# Estimate mass increase

# Split data by species in new database
deli_stressor_2 <- data_2 %>% filter (species == "delicata")
guich_stressor_2 <- data_2 %>% filter (species == "guichenoti")

# Explore distribution of mass increase
  # A) L.deli
plot_mass <- ggplot(deli_stressor_2, aes(x = mass, fill = interaction(as.factor(temp),as.factor(cort), sep="_"))) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 30, color = "black") +
  scale_fill_manual(values = c("28_A" = "red", "23_A" = "blue", "28_B" = "salmon", "23_B" = "lightblue")) +
  labs(x = "Mass", y = "Density", fill = "Temperature ~ CORT") +
  theme_minimal() +
  facet_grid(cort*temp~Day, scales = "free_y")
print(plot_mass)

mass_normal_test<-by(deli_stressor_2$mass, interaction (deli_stressor$cort, deli_stressor$temp, deli_stressor$Day), function(x) shapiro.test((x)))
print(mass_normal_test)

  #B) L.guich
plot_mass <- ggplot(guich_stressor_2, aes(x = mass, fill = interaction(as.factor(temp),as.factor(cort), sep="_"))) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 30, color = "black") +
  scale_fill_manual(values = c("28_A" = "red", "23_A" = "blue", "28_B" = "salmon", "23_B" = "lightblue")) +
  labs(x = "Mass", y = "Density", fill = "Temperature ~ CORT") +
  theme_minimal() +
  facet_grid(cort*temp~Day, scales = "free_y")
print(plot_mass)

mass_normal_test<-by(guich_stressor$mass, interaction (guich_stressor$cort, guich_stressor$temp, guich_stressor$Day), function(x) shapiro.test((x)))
print(mass_normal_test)

# BEHAVIOUR
# Split data by species in original database
deli_stressor <- data %>% filter (species == "delicata")
guich_stressor <- data %>% filter (species == "guichenoti")
# Explore distribution of latency to move/time frozen
  # A) L. deli
plot_freeze <- ggplot(deli_stressor, aes(x = log(lat_to_move+1), fill = interaction(as.factor(temp),as.factor(cort), sep="_"))) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 30, color = "black") +
  scale_fill_manual(values = c("28_A" = "red", "23_A" = "blue", "28_B" = "salmon", "23_B" = "lightblue")) +
  labs(x = "Mass", y = "Density", fill = "Temperature ~ CORT") +
  theme_minimal() +
  facet_grid(cort*temp~Day, scales = "free_y")
print(plot_freeze)

freeze_normal_test<-by(deli_stressor$lat_to_move, interaction (deli_stressor$cort, deli_stressor$temp, deli_stressor$Day), function(x) shapiro.test(log(x+1)))
print(freeze_normal_test)

 #B) L.guich
plot_freeze <- ggplot(guich_stressor, aes(x = log(lat_to_move+1), fill = interaction(as.factor(temp),as.factor(cort), sep="_"))) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 30, color = "black") +
  scale_fill_manual(values = c("28_A" = "red", "23_A" = "blue", "28_B" = "salmon", "23_B" = "lightblue")) +
  labs(x = "Mass", y = "Density", fill = "Temperature ~ CORT") +
  theme_minimal() +
  facet_grid(cort*temp~Day, scales = "free_y")
print(plot_freeze)

freeze_normal_test<-by(guich_stressor$lat_to_move, interaction (guich_stressor$cort, guich_stressor$temp, guich_stressor$Day), function(x) shapiro.test(log(x+1)))
print(freeze_normal_test)
